---
title: "431 Class 18"
author: Thomas E. Love, Ph.D.
date: "2024-10-29"
format:
  revealjs: 
    theme: default
    self-contained: true
    slide-number: true
    footnotes-hover: true
    preview-links: auto
    date-format: iso
    logo: 431_2024_logo.png
    footer: "431 Class 18 | 2024-10-29 | <https://thomaselove.github.io/431-2024/>"
---

## Today's Agenda

Two-Factor (Two-Way) Analysis of Variance

- The importance of interaction
- Interpreting ANOVA output
- What are the key model checks for ANOVA?
- Comparing Model Performance Indices
- What if we add a covariate?

## R Packages

```{r}
#| echo: true

library(janitor)
library(readxl)
library(easystats)
library(tidyverse)

source("c18/data/Love-431.R")
knitr::opts_chunk$set(comment = NA)

theme_set(theme_bw())
```

## The dm464 data, in full

We've discussed subsets of these data in previous classes.

```{r}
#| echo: true
dm464 <- read_xlsx("c18/data/dm464.xlsx") |>
  janitor::clean_names() |>
  mutate(across(where(is.character), as_factor)) |>
  mutate(id_code = as.character(id_code),
         smoke_start = fct_relevel(smoke_start, "Current", "Former"),
         smoke_end = fct_relevel(smoke_end, "Current", "Former"),
         insur_start = fct_relevel(insur_start, "Medicare", 
                                   "Commercial", "Medicaid"),
         insur_end = fct_relevel(insur_end, "Medicare", 
                                   "Commercial", "Medicaid"))

dim(dm464)
```

## Descriptions for Today's Variables {.smaller}

-   Cohort of 464 English-speaking adults with diabetes measured at the start and end of a two-year period.

| Variable | Description |
|---------------:|:-------------------------------------------------------|
| `id_code` | DM-xxxx code to identify subjects (not consecutive) |
| `bmi_end` | Body Mass Index in $(kg/m^2)$ at the end of the period |
| `dep_start` | Depression diagnosis at the start of the period (1 = yes, 0 = no) |
| `ed_start` | High = more than 78% of adult residents of subject's home neighborhood graduated high school, vs. Low = no more than 78%. |
| `insur_start` | Subject's insurance status at the start (3 levels) |

-   78% was close to the median value of the quantitative measure of high school graduation rate among enrolled subjects at the start of the period.
-   All subjects have either Medicare, Commercial or Medicaid insurance.

## Today's Variables

```{r}
#| echo: true
dm18 <- dm464 |>
  mutate(ed_start = factor(ifelse(nedlev_start > 78, "High", "Low")),
         dep_start = factor(depdiag_start)) |>
  select(id_code, bmi_end, dep_start, ed_start, insur_start)

head(dm18)
```

## `data_codebook()` results

```{r}
#| echo: true
data_codebook(dm18 |> select (-id_code))
```

# Two-Factor Analysis of Variance: Example 1

## Example 1

Suppose we want to simultaneously understand the impacts of two factors on BMI at the end of the period, specifically:

-   `dep_start` (1 or 0), and
-   `ed_start` (Low or High)

and it is possible that the impact of depression diagnosis on BMI may depend on the home neighborhood's educational attainment, and vice versa (i.e. depression and education may **interact**.)

## An interaction plot is a plot of means.

Calculate and store the group means.

```{r}
#| echo: true
ex1_means <- dm18 |>
  group_by(dep_start, ed_start) |>
  summarize(mean_bmi = mean(bmi_end))

ex1_means
```

## How strong is the interaction?

Now, we'll plot these means, and look for a substantial interaction (non-parallel lines.)

```{r}
#| echo: true
#| output-location: slide

ggplot(ex1_means, aes(x = dep_start, y = mean_bmi)) +
  geom_line(aes(group = ed_start, color = ed_start)) +
  geom_point(aes(color = ed_start)) +
  scale_color_material() +
  labs(title = "Example 1 Interaction Plot",
       subtitle = "Do we see substantially non-parallel lines?",
       y = "Mean BMI at end of period", 
       x = "Depression Diagnosis (1 = yes, 0 = no)", 
       color = "HS Graduation Rate")
```

## Two-Way ANOVA without interaction

This model predicts our outcome (`bmi_end`) using the two factors as main effects only (with no interaction.)

```{r}
#| echo: true

mainfx <- lm(bmi_end ~ dep_start + ed_start, data = dm18)

mainfx
```

$$
bmi_{end} = 32.98 + 3.452 (dep_{start} = 1) + 1.555 (ed_{start} = Low) 
$$

## Interpret the estimate: 3.45?

```{r}
#| echo: true
model_parameters(mainfx, ci = 0.90)
```

> -   If we have two subjects who live in neighborhoods with the same `ed_start` value, but only one has a depression diagnosis at the start of the period, then **on average** the subject with a depression diagnosis will have a BMI that is 3.45 $kg/m^2$ higher than the subject without such a diagnosis, according to our `mainfx` model.

## Interpret the estimate: 1.56?

```{r}
#| echo: true
model_parameters(mainfx, ci = 0.90)
```

> -   Our `mainfx` model estimates that if we have two subjects who have the same depression diagnosis status at the start of the period, but live in neighborhoods with different `ed_start` values (one High and one Low), then **on average** the subject living in the Low education neighborhood will have a BMI that is 1.56 $kg/m^2$ higher than the other subject.

## Interpret the intercept term?

```{r}
#| echo: true
model_parameters(mainfx, ci = 0.90)
```

> -   32.98 is the intercept. It's the `mainfx` model's estimate of the mean of `bmi_end` across all subjects where both `dep_start` is 0 and `ed_start` = High.

## ANOVA table for main effects model

```{r}
#| echo: true
anova(mainfx)
```

### % of Total SS explained by the two factors?

-   Total SS = 1303 + 280 + 35113 = 36696
-   So % explained = (1303 + 280) / 36696 = 0.043, or 4.3%

## Model `mainfx` and its assumptions

-   `mainfx` is a "main effects only" model. It assumes there is zero interaction between `dep_start` and `ed_start` in predicting our outcome, `bmi_end`.
-   Like any ANOVA model, `mainfx` makes the usual linear model assumptions (linearity, constant variance, Normality)
    -   Focus: posterior predictive check, and Normal Q-Q plot of residuals[^1].

[^1]: The assessment of linearity and non-constant variance used by `check_model()` are best used when we have at least one quantitative predictor in our model.

## Checking model `mainfx`

```{r}
#| echo: true
set.seed(20241029)
check_model(mainfx, check = c("pp_check", "qq"), detrend = FALSE)
```

## Interaction Plot, again

```{r}
#| echo: false

ggplot(ex1_means, aes(x = dep_start, y = mean_bmi)) +
  geom_line(aes(group = ed_start, color = ed_start)) +
  geom_point(aes(color = ed_start)) +
  scale_color_material() +
  labs(title = "Example 1 Interaction Plot",
       subtitle = "Do we see substantially non-parallel lines?",
       y = "Mean BMI at end of period", 
       x = "Depression Diagnosis (1 = yes, 0 = no)", 
       color = "HS Graduation Rate")
```

## Two-Way ANOVA with Interaction

Fit the linear model, including an interaction (`*`) between our factors.

```{r}
#| echo: true

withint <- lm(bmi_end ~ dep_start * ed_start, data = dm18)

withint
```

## What is the `withint` model equation?

$$
bmi_{end} = 33.4 + 2.3 (dep_{start} = 1) \\
+ 0.7 (ed_{start} = Low) \\ + 2.2 (dep_{start} = 1) \times (ed_{start} = Low)
$$

`dep_start` | `ed_start` | Estimated `bmi_end`
:---------: | :--------: | -------------:
0 | High | 33.4  
0 | Low | 33.4 + 0.7 = 34.1
1 | High | 33.4 + 2.3 = 35.7
1 | Low | 33.4 + 2.3 + 0.7 + 2.2 = 38.6

## Effect of High vs. Low `ed_start`?

Est. `bmi_end` | `ed_start` = High | `ed_start` = Low
:---------: | :--------: | :----------:
`dep_start` = 0 | 33.4 | 34.1  
`dep_start` = 1 | 35.7 | 38.6

- If `dep_start` = 0, High `ed_start` yields 33.4, and Low yields 34.1, so the High - Low difference in estimates is -0.7.
- If `dep_start` = 1, High `ed_start` yields 35.7, and Low yields 38.6, so the High - Low difference in estimates is -2.9.
- Estimated effect of `ed_start` **depends** on `dep_start`.

## Effect of `dep_start` = 1 vs. 0?

Est. `bmi_end` | `ed_start` = High | `ed_start` = Low
:---------: | :--------: | :----------:
`dep_start` = 1 | 35.7 | 38.6
`dep_start` = 0 | 33.4 | 34.1  

- If `ed_start` is High, then `dep_start` = 1 yields 35.7, and 0 yields 33.4, so the difference in estimated `bmi_end` is 2.3.
- If `ed_start` is High, then `dep_start` = 1 yields 38.6, and 0 yields 34.1, so the difference in estimated `bmi_end` is 4.5.
- Estimated effect of `dep_start` **depends** on `ed_start`.

## ANOVA table for `withint` model

```{r}
anova(withint)
```

- Total SS = 1303 + 280 + 133 + 34979 = 36696
- % explained = (1303 + 280 + 133) / 36696 = 0.047, or 4.7%
- Is this a big improvement over the `mainfx` model?

## Checking model `withint`

```{r}
#| echo: true
set.seed(202410291)
check_model(withint, check = c("pp_check", "qq"), detrend = FALSE)
```

## Compare peformance of our two models?

```{r}
#| echo: true

compare_performance(mainfx, withint, rank = TRUE)
```

## Plot to compare performance indicators

```{r}
#| echo: true
plot(compare_performance(mainfx, withint))
```

## Do we have a substantial interaction?

- Interaction plot shows slightly non-parallel lines.
- % of variation explained by the interaction term is about 0.4% of the variation in `bmi_end`, overall.
- p value for the interaction term is pretty large ($p$ = 0.19)
- Is there a substantial improvement in the fit with the interaction?
    - In-sample performance indices are split.

What do you think?

# Two-Factor Analysis of Variance: Example 2

## Example 2

Suppose that we now want to understand the impact on BMI at the end of the period, of:

- `insur_start` (Medicare, Commercial or Medicaid), and
- `ed_start` (Low or High)

and it's possible that the impact of insurance on BMI may depend on the home neighborhood's educational attainment, and vice versa (i.e. insurance and education may **interact**.)

## Build an interaction plot

Calculate and store group means.

```{r}
#| echo: true
ex2_means <- dm18 |>
  group_by(insur_start, ed_start) |>
  summarize(mean_bmi = mean(bmi_end))

ex2_means
```

## How strong is the interaction?

Now, we'll plot these means, and look for a substantial interaction (non-parallel lines.)

```{r}
#| echo: true
#| output-location: slide

ggplot(ex2_means, aes(x = insur_start, y = mean_bmi)) +
  geom_line(aes(group = ed_start, color = ed_start)) +
  geom_point(aes(color = ed_start)) +
  scale_color_material() +
  labs(title = "Example 2 Interaction Plot",
       subtitle = "Do we see substantially non-parallel lines?",
       y = "Mean BMI at end of period", 
       x = "Insurance at start of period", 
       color = "HS Graduation Rate")
```

## Example 2: Two-Way ANOVA models 

```{r}
#| echo: true

mainfx2 <- lm(bmi_end ~ insur_start + ed_start, data = dm18)
withint2 <- lm(bmi_end ~ insur_start * ed_start, data = dm18)

compare_performance(mainfx2, withint2, rank = TRUE)
```

## Plotting Performance Indices

```{r}
#| echo: true

plot(compare_performance(mainfx2, withint2))
```

## Interpret `mainfx2` estimates?

```{r}
#| echo: true
model_parameters(mainfx2, ci = 0.90)
```

$$
bmi_{end} = 32.9 + 1.4 (Comm.) + 2.4 (Medicaid) + 1.8 (Low Ed.) 
$$

Est. `bmi_end` | Medicare | Commercial | Medicaid
------------: | :------: | :------: | :------: 
`ed_start` High | 32.9 | 34.3 | 35.3 
`ed_start` Low | 34.7 | 36.1 | 37.1


## Specify `withint2` equation?

```{r}
#| echo: true
model_parameters(withint2, ci = 0.90)
```

## `withint2` model equation {.smaller}

$$
bmi_{end} = 33.6 + 2.3 Commercial + 0.7 Medicaid + 0.5 (ed_{start} = Low) \\ - 1.9 Commercial \times (ed_{start} = Low) + 3.4 Medicaid \times (ed_{start} = Low)
$$

`insur_start` | `ed_start` | Estimated `bmi_end`
:---------: | :--------: | -------------:
Medicare | High | 33.6
Commercial | High | 33.6 + 2.3 = 35.9
Medicaid | High | 33.6 + 0.7 = 34.3
Medicare | Low | 33.6 + 0.5 = 34.1
Commercial | Low | 33.6 + 2.3 + 0.5 - 1.9 = 34.5
Medicaid | Low | 33.6 + 0.7 + 0.5 + 3.4 = 38.2

## ANOVA table for main effects model

```{r}
#| echo: true
anova(mainfx2)
```

### % of Total SS explained by the two factors?

-   Total SS = 510 + 369 + 35817 = 36696
-   So % explained = (510 + 369) / 36696 = 0.024, or 2.4%

## ANOVA table for interaction model

```{r}
#| echo: true
anova(withint2)
```

### % of Total SS explained by the model?

-   Total SS = 510 + 369 + 482 + 35335 = 36696
-   So % explained = (510 + 369 + 482) / 36696 = 0.037, or 3.7%

## Checking model `mainfx2`

```{r}
#| echo: true
set.seed(202410292)
check_model(mainfx2, check = c("pp_check", "qq"), detrend = FALSE)
```

## Checking model `withint2`

```{r}
#| echo: true
set.seed(202410293)
check_model(withint2, check = c("pp_check", "qq"), detrend = FALSE)
```

## Do we have a substantial interaction?

Example 2

- Interaction plot shows substantially non-parallel lines.
- Interaction term explains about 1.3% of the variation in `bmi_end`, overall.
- p value for the interaction term is fairly small ($p$ = 0.045)
- Substantial improvement in the fit with interaction?
    - 6/7 in-sample performance indices prefer interaction

What do you think?

# Add a covariate?

## What if we added a covariate?

```{r}
#| echo: true
dm18a <- dm464 |>
  mutate(ed_start = factor(ifelse(nedlev_start > 78, "High", "Low")),
         dep_start = factor(depdiag_start)) |>
  mutate(bmi_c = center(bmi_start)) |>
  select(id_code, bmi_end, dep_start, ed_start, insur_start, 
         bmi_c, bmi_start)

dm18a |> reframe(lovedist(bmi_start))
dm18a |> reframe(lovedist(bmi_c))
```

## Analysis of Covariance Model

```{r}
#| echo: true

fit3 <- lm(bmi_end ~ bmi_c + dep_start * insur_start, data = dm18a)

anova(fit3)
```

## Proportion of Variation Explained

```{r}
#| echo: true

eta_squared(fit3)
```

## Drop Interaction?

```{r}
#| echo: true

fit4 <- lm(bmi_end ~ bmi_c + dep_start + insur_start, data = dm18a)

anova(fit4)
```

## Drop Depression?

```{r}
#| echo: true

fit5 <- lm(bmi_end ~ bmi_c + insur_start, data = dm18a)

anova(fit5)
```

## Compare fit3, fit4, fit5

```{r}
compare_performance(fit3, fit4, fit5, rank = TRUE)
```

## Compare fit3, fit4, fit5

```{r}
plot(compare_performance(fit3, fit4, fit5))
```

## Check Model `fit4` (1/3)

```{r}
#| echo: true
check_model(fit4, check = c("pp_check", "qq"), detrend = FALSE)
```

## Check Model `fit4` (2/3)

```{r}
#| echo: true
check_model(fit4, check = c("linearity", "homogeneity"))
```

## Check Model `fit4` (3/3)

```{r}
#| echo: true
check_model(fit4, check = c("outliers", "vif"))
```

## Session Information

```{r}
#| echo: true
xfun::session_info()
```
